---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Setup
The `tidyverse` makes this package easy to use. The golden ratio sets up the correct proportions for several measurements. So, we'll save it as a global variable for this vignette.

```{r setup, message = FALSE}
library(penrosetiling)
library(tidyverse)

g_ratio <- (1 + sqrt(5)) / 2
```

## Background

This package uses Robinson triangles to subdivide the rhombi in Penrose tiling. These triangles have specific proportions and the subdivision follows a set of rules. In this first example, the four types of triangles are displayed. Regardless of how they are actually positioned, the triangle type assumes these isosceles triangles sit flat with the two equal sides pointing upwards. There are two shapes, "t" for thin and "T" for thick, and two directions, left and right. Left triangles have their first point on the left side of the middle point. Right triangles are flipped. For the functions in this package, data sets with triangles will need the following columns: `triangle` for triangle id, `rhombus` for rhombus id, `point1_x`, `point1_y`, `point2_x`, `point2_y`, `point3_x`, and `point3_y` for coordinates.

The following code creates an introduction data set with each of the four triangles. The `pivot_down` function drops the data set from one row per triangle to one row per point with a new `point` column.

```{r background_1, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- data.frame(triangle = seq(1, 4),
                    shape = c("t", "t", "T", "T"),
                    point1_x = c(0, 2.5, 3, 5.5 + 2),
                    point1_y = c(0, 0, 0, 0),
                    point2_x = c(g_ratio * cos(72 * pi/180),
                                 1.5 + g_ratio * cos(72 * pi/180),
                                 3 + 2 * (1/g_ratio) * cos(36 * pi/180),
                                 5.5 + 2 * (1/g_ratio) * cos(36 * pi/180)),
                    point2_y = c(g_ratio * sin(72 * pi/180),
                                 g_ratio * sin(72 * pi/180),
                                 2 * (1/g_ratio) * sin(36 * pi/180),
                                 2 * (1/g_ratio) * sin(36 * pi/180)),
                    point3_x = c(1, 1.5, 3 + 2, 5.5),
                    point3_y = c(0, 0, 0, 0),
                    rhombus = seq(1, 4))

tiles_start <- pivot_down(tiles) %>%
  group_by(triangle) %>%
  mutate(mean_x = mean(x),
         mean_y = mean(y)) %>%
  mutate(label_x = .66 * x + .34 * mean_x,
         label_y = .66 * y + .34 * mean_y)

ggplot(tiles_start) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  geom_text(aes(x = label_x,
                y = label_y,
                label = point)) +
  scale_fill_manual(values = c("#FFC749",
                               "#B0E2FF")) +
  coord_equal() +
  theme_void() +
  theme(legend.position = "bottom")
```

The four triangles are displayed with each of their points labels. The left triangles have corner 1 to the left of the corner pointing upwards. If these triangles were rotated upside down, the point 1 and 3 would not flip sides and the triangles' direction would stay the same.. Left triangles always have point 1 to the left of point 2 when viewing the triangle from edge connecting points 1 and 3.

The `substitution` function subdivides each triangle appropriately, depending on shape and direction. It will return a larger data set of smaller triangles. The smaller triangles have specific shapes and directions based on the input triangles.

```{r background_2, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles_substitution <- substitution(tiles)

tiles_substitution <- pivot_down(tiles_substitution) %>%
  group_by(triangle) %>%
  mutate(mean_x = mean(x),
         mean_y = mean(y)) %>%
  mutate(label_x = .66 * x + .34 * mean_x,
         label_y = .66 * y + .34 * mean_y)

ggplot(tiles_substitution) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  geom_text(aes(x = label_x,
                y = label_y,
                label = point)) +
  scale_fill_manual(values = c("#FFC749",
                               "#B0E2FF")) +
  coord_equal() +
  theme_void() +
  theme(legend.position = "bottom")
```

The first triangle was thin and left. The `substitution` subdivided it into two smaller left triangles, one thin and one thick. The thick triangles are converted into three smaller ones. 

## Starting Data Sets

There are two functions that provide data sets to start coding.

The `return_rhombus` function takes a starting position and shape to create a data set with two observations: two triangles that make up a rhombus.

```{r return_rhombus_1, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles_1 <- return_rhombus(0, 0, -90 + 36, 2, 'T') %>%
  pivot_down() %>%
  group_by(triangle) %>%
  mutate(level = 1)

tiles_2 <- return_rhombus(0, 0, -90 + 36, 2, 'T') %>%
  substitution() %>%
  pivot_down() %>%
  group_by(triangle) %>%
  mutate(level = 2)

tiles_5 <- return_rhombus(0, 0, -90 + 36, 2, 'T') %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 5)

tiles <- bind_rows(tiles_1, tiles_2, tiles_5)
  
ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal() +
  facet_wrap(~ level)
```

```{r return_rhombus_2, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles_1 <- return_rhombus(0, 0, 0, 1 / sin(72 * pi/180), 't') %>%
  pivot_down() %>%
  group_by(triangle) %>%
  mutate(level = 1)

tiles_2 <- return_rhombus(0, 0, 0, 1 / sin(72 * pi/180), 't') %>%
  substitution() %>%
  pivot_down() %>%
  group_by(triangle) %>%
  mutate(level = 2)

tiles_5 <- return_rhombus(0, 0, 0, 1 / sin(72 * pi/180), 't') %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 5)

tiles <- bind_rows(tiles_1, tiles_2, tiles_5)

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal() +
  facet_wrap(~ level)
```

The `return_starting_circle` function also takes a starting position and shape to create a data set. This time the data set has ten observations. Either five rhombi or ten half rhombi depending on the type of triangle.

```{r return_starting_circle_1, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles_1 <- return_starting_circle(0, 0, 0, 10, 't') %>%
  pivot_down() %>%
  mutate(level = 1)

tiles_2 <- return_starting_circle(0, 0, 0, 10, 't') %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 2)

tiles_5 <- return_starting_circle(0, 0, 0, 10, 't') %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 5)

tiles <- bind_rows(tiles_1, tiles_2, tiles_5)

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal() +
  facet_wrap(~ level)
```


```{r return_starting_circle_2, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles_1 <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  pivot_down() %>%
  mutate(level = 1)

tiles_2 <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 2)

tiles_5 <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  pivot_down() %>%
  mutate(level = 5)

tiles <- bind_rows(tiles_1, tiles_2, tiles_5)

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal() +
  facet_wrap(~ level)
```
