---
title: "Using Purrr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-purrr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

```{r setup, message = FALSE}
library(penrosetiling)
library(tidyverse)
```

## Basic Repetition

You can subdivide the same data down by calling `substitution` multiple times.
```{r repetition_1, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- return_rhombus(0, 0, -90 + 36, 100, 'T') %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  substitution() %>%
  pivot_down()

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal()
```
</br>
You can also use a loop.
```{r repetition_2, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- return_rhombus(0, 0, -90 + 36, 100, 'T') 

for(i in seq(1, 5)) {
  tiles <- tiles %>% substitution()
}
  
tiles <- tiles %>%
  pivot_down()

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal()
```
</br>
Using `purrr::reduce` will wrap everything up nicely into a function call. You'll need to pass a function with two parameters for it to work. This makes it easy to adjust the number of substitutions.
```{r repetition_3, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- return_rhombus(0, 0, -90 + 36, 100, 'T')

tiles <- purrr::reduce(.x = rep(list(1), 5),
                       .f = function(df, x) {substitution(df)},
                       .init = tiles)

tiles <- tiles %>%
  pivot_down()

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal()
```

## Advance Repetition

This setup allows for passing a function that subdivides different sections of the data set. For example, the number of times a triangle is subdivided can be a function of the mean x position.

```{r x_position, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- return_rhombus(0, 0, -90 + 36, 100, 'T') %>%
  substitution() %>%
  substitution() %>%
  pivot_down() %>%
  group_by(triangle) %>%
  mutate(mean_x = mean(x))

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = mean_x,
                   group = triangle)) +
  coord_equal()
```

By creating a function that checks if the x position is past a cutoff, then passing the cutoffs as a list, `purrr:reduce` will loop through the cutoffs and apply the function appropriately. In this example, triangles with a mean x position past the cutoffs of 10, 20, ..., 80 will be subdivided.

```{r x_cutoffs, fig.align = 'center', fig.height = 2, fig.width = 7}
check_substitution <- function(df, x) {
  df <- df %>%
    rowwise() %>%
    mutate(mean_x = c(point1_x + point2_x + point3_x) / 3) %>%
    mutate(subdivide = if_else(mean_x > x, TRUE, FALSE))
  
  df_subdivide <- df %>%
    filter(subdivide == TRUE) %>%
    select(-subdivide) %>%
    substitution()
  
  df <- df %>%
    filter(subdivide == FALSE) %>%
    select(-subdivide) %>%
    bind_rows(df_subdivide) %>%
    select(-c(triangle, rhombus)) %>% # reset triangle and rhombus
    rowid_to_column("triangle") %>%
    group_by(point1_x, point1_y, point3_x, point3_y) %>%
    mutate(rhombus = cur_group_id()) %>%
    ungroup()
  
  return(df)
}

tiles <- return_rhombus(0, 0, -90 + 36, 100, 'T')

tiles <- purrr::reduce(.x = as.list(seq(10, 80, by = 10)),
                .f = check_substitution,
                .init = tiles)

tiles <- tiles %>%
  pivot_down()

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = shape,
                   group = triangle)) +
  coord_equal()
```
</br>
In the next example, the cutoffs will be based on the distance to the center. This will also use the center of the rhombus instead of the triangle.
```{r center_cutoffs, fig.align = 'center', fig.height = 2, fig.width = 7}
tiles <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  substitution() %>%
  pivot_down() %>%
  select(-triangle) %>%
  distinct() %>%
  group_by(rhombus) %>%
  mutate(center_x = mean(x), 
         center_y = mean(y)) %>%
  mutate(distance = round(sqrt(center_x^2 + center_y^2), 2))
  
ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = as.character(distance),
                   group = rhombus)) +
  coord_equal() +
  scale_fill_discrete("Distance from center") +
  theme(legend.position = "bottom")
```
</br>
We'll use cutoffs of 3 and 5. This will subdivide the middle and outside once, then the outside another time. Now we just need to stick this in a function, subdivide the ones that pass, and join back to the triangles that didn't.
```{r center_cutoffs_1, fig.align = 'center', fig.height = 2, fig.width = 7}
check_substitution <- function(df, x) {
  df <- df %>%
    mutate(mean_x = c(point1_x + point2_x + point3_x) / 3,
           mean_y = c(point1_y + point2_y + point3_y) / 3) %>%
    group_by(rhombus) %>%
    mutate(mean_x = mean(mean_x),
           mean_y = mean(mean_y)) %>%
    ungroup() %>%
    mutate(distance = sqrt(mean_x^2 + mean_y^2)) %>%
    mutate(subdivide = if_else(distance > x, TRUE, FALSE)) 
  
  df_subdivide <- df %>%
    filter(subdivide == TRUE) %>%
    select(-subdivide) %>%
    substitution() %>%
    substitution()
  
  df <- df %>%
    filter(subdivide == FALSE) %>%
    select(-subdivide) %>%
    bind_rows(df_subdivide) %>%
    select(-c(triangle, rhombus)) %>% # reset triangle and rhombus
    rowid_to_column("triangle") %>%
    group_by(point1_x, point1_y, point3_x, point3_y) %>%
    mutate(rhombus = cur_group_id()) %>%
    ungroup()
  
  return(df)
}

tiles <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  substitution()

tiles <- purrr::reduce(.x = as.list(c(3, 5)),
                       .f = check_substitution,
                       .init = tiles)

tiles <- tiles %>%
  pivot_down() %>%
  select(-triangle) %>%
  distinct() %>%
  group_by(rhombus) %>%
  mutate(center_x = mean(x), 
         center_y = mean(y)) %>%
  mutate(distance = round(sqrt(center_x^2 + center_y^2), 2))

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = distance,
                   group = rhombus)) +
  coord_equal()
```
</br>
A small catch here is that the rhombus centers change after subdivision. If you want to keep the same distances, just keep the centers from the beginning and don't recalculate.
```{r center_cutoffs_2, fig.align = 'center', fig.height = 2, fig.width = 7}
check_substitution <- function(df, x) {
  df <- df %>%
    mutate(subdivide = if_else(distance > x, TRUE, FALSE)) 
  
  df_subdivide <- df %>%
    filter(subdivide == TRUE) %>%
    select(-subdivide) %>%
    substitution() %>%
    substitution()
  
  df <- df %>%
    filter(subdivide == FALSE) %>%
    select(-subdivide) %>%
    bind_rows(df_subdivide) %>%
    select(-c(triangle, rhombus)) %>% # reset triangle and rhombus
    rowid_to_column("triangle") %>%
    group_by(point1_x, point1_y, point3_x, point3_y) %>%
    mutate(rhombus = cur_group_id()) %>%
    ungroup()
  
  return(df)
}

tiles <- return_starting_circle(0, 0, 90, 5, 'T') %>%
  substitution() %>%
  mutate(mean_x = c(point1_x + point2_x + point3_x) / 3,
         mean_y = c(point1_y + point2_y + point3_y) / 3) %>%
  group_by(rhombus) %>%
  mutate(mean_x = mean(mean_x),
         mean_y = mean(mean_y)) %>%
  ungroup() %>%
  mutate(distance = sqrt(mean_x^2 + mean_y^2)) 

tiles <- purrr::reduce(.x = as.list(c(3, 5)),
                       .f = check_substitution,
                       .init = tiles)

tiles <- tiles %>%
  pivot_down() %>%
  select(-triangle) %>%
  distinct() %>%
  group_by(rhombus) %>%
  mutate(center_x = mean(x), 
         center_y = mean(y)) %>%
  mutate(distance = round(sqrt(center_x^2 + center_y^2), 2))

ggplot(tiles) +
  geom_polygon(aes(x = x,
                   y = y,
                   fill = distance,
                   group = rhombus)) +
  coord_equal()
```
There are slight differences between this image and the previous one.
